---
title: "College Football Performance Analysis"
output: html_document
date: "2024-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## FPI Data Scraping
```{r}
library(dplyr)
library(rvest)
library(stringr)
library(purrr)

scrape_fpi <- function(season = 2020, stat = "FPI"){
  
  current_year <- str_sub(Sys.Date(), 1, 4) %>% as.integer()
  
  if (between(season, 2012, current_year) == FALSE)
  
  if (!(stat %in% c("FPI", "RESUME", "EFF")) == TRUE)
  
  message(glue::glue("Scraping {stat} for {season}"))
  
  url_fpi <- glue::glue("https://www.espn.com/college-football/fpi/_/view/fpi/season/{season}")
  url_resume <- glue::glue("https://www.espn.com/college-football/fpi/_/view/resume/season/{season}")
  url_eff <- glue::glue("https://www.espn.com/college-football/fpi/_/view/efficiencies/season/{season}")
  
  if (stat == "FPI"){
    fpi_html <- read_html(url_fpi)
    
    fpi_names <- c(
      "team", "conf","w_l", "fpi", "rk", "trend","proj", "proj2", 
      "proj3", "proj4", "proj5", "proj6", "proj7", "proj8"
    )
    
    trend_data <- fpi_html %>%
      html_nodes("td:nth-child(4) > div") %>%
      html_attr("class")
    
    table_fpi <- suppressWarnings(
      fpi_html %>%
        html_table() %>%
        cbind.data.frame() %>%
        set_names(nm = fpi_names) %>%
        slice(-1) %>%
        mutate(
          across(c(fpi), as.double),
          across(c(rk, trend), as.integer),
          season = season
        ) %>%
        mutate(
          trend = if_else(
            trend_data[rk] == "trend negative",
            as.integer(trend * -1),
            trend)
        ) %>%
        select(season, team, conf, fpi, rk) %>%
        tibble()
    )
    
    table_fpi
    
  } else if (stat == "RESUME"){
    resume_html <- read_html(url_resume)
    
    resume_names <- c(
      "team", "conf", "sor_rk", "fpi_rk", "ap_cfp_rk", "sos_rk",
      "rem_sos_rk", "gc_rk", "avgwp_rk"
    )
    
    table_resume <- suppressWarnings(
      resume_html %>%
      html_table() %>%
      cbind.data.frame()  %>%
      set_names(nm = resume_names) %>%
      slice(-1)%>%
      mutate(
          across(c(sor_rk:avgwp_rk), as.integer),
          season = season
        ) %>%
      select(season, everything()) %>%
      tibble()
    
    ) 
      table_resume
    
    
  } else if (stat == "EFF"){
    eff_html <- read_html(url_eff)
    
    eff_names <- c(
      "team", "conf", "eff_ove", "rnk_ove", "eff_off", "rnk_off",
      "eff_def", "rnk_def", "eff_spe", "rnk_spe"
    )
    
    table_eff <-
      eff_html %>%
      html_table() %>%
      cbind.data.frame() %>%
      set_names(nm=eff_names) %>%
      slice(-1) %>%
      mutate(
        across(contains("rnk"), as.integer),
        across(contains("eff"), as.double),
        season = season
      ) %>%
      select(season, everything()) %>%
      tibble()
    
    table_eff
  }
  
}
```

## Exploratory Data Analysis

```{r}
library(ggplot2)
library(scales)
library(gridExtra)
library(tidyverse)
library(dplyr)

expenses <- c( "Other_Expenses", "Excess_Transfers_Back", "Medical", 
               "Competition_Guarantees_spent", "Recruiting", "Game_Expenses_and_Travel", 
               "Facilities_and_Equipment", "Coaches_Compensation", 
               "Support_and_Admin_Compensation_w_Severance", "Athletic_Student_Aid", 
               "Total_Football_Spending", "Total_Football_Coaching_Salaries" )
revenues <- c( "Corporate_Sponsorship_Advertising_Licensing", 
               "Donor_Contributions", "Competition_Guarantees_earned", 
               "NCAA_Conference_Distributions_Media_Rights_and_Post_Season_Football", 
               "Ticket_Sales", "Institutional_Government_Support")

metrics <- c("rk", "eff_ove", "rnk_ove", "eff_off", "rnk_off", "eff_def",
             "rnk_def", "eff_spe", "rnk_spe")

df_fpi_year_off <- read_csv("df_fpi_year_off.csv") %>%
  select(Data, IPEDS_ID, year, fpi, Total_Expenses, all_of(expenses), Total_Revenues, all_of(revenues)) %>%
  mutate(ExpRev = Total_Expenses / Total_Revenues,
         FBExp = Total_Football_Spending / Total_Expenses)
         


# Conference key for looping plots
Conference_key <- read_csv("Conference_key.csv") 

# Original Data
df <- read_csv("df_03202024.csv") %>%
  select(c("Data", "IPEDS_ID", "year", "fpi", "Total_Expenses", all_of(expenses), "Total_Revenues", all_of(revenues))) %>%
  mutate(ExpRev = Total_Expenses / Total_Revenues,
         FBExp = Total_Football_Spending / Total_Expenses)

# difference in FPI
df_delta <- read.csv("deltas_03202024.csv", header = TRUE, row.names = NULL) %>%
  select(c("Data", "IPEDS_ID", "year", "delta_fpi", "Total_Expenses", all_of(expenses), "Total_Revenues", all_of(revenues))) %>%
  mutate(ExpRev = Total_Expenses / Total_Revenues,
         FBExp = Total_Football_Spending / Total_Expenses)

```

### Summary Statistics
```{r}
# Boxplot of numerical variables
df_numeric <- df %>%
  select(-c(Data, IPEDS_ID, year, fpi, ExpRev, FBExp)) # Exclude non-numeric variables

df_numeric_long <- df_numeric %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

ggplot(df_numeric_long, aes(x = Variable, y = Value)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  coord_flip() +
  scale_y_continuous(trans = "log10") +
  labs(title = "Boxplot of Monetary Variables (Log scale)", y = "$ (log scale)")

# Expenditure of NCAA over time
# find total expenses grouped by year
yearly<-df%>%
  group_by(year)%>%
  summarize(Total_Rev = sum(Total_Revenues,na.rm=TRUE)/1000000000,
            Total_Exp=sum(Total_Expenses,na.rm=TRUE)/1000000000)

# get all years except for 2023 (due to no data for that year)
yearly<-yearly[yearly$year!=2023,]

# line plot of the expenses over time
ggplot(yearly, aes(x=year, y=Total_Exp))+geom_point()+geom_line()+theme_minimal()+labs(x="Year",y="Total Expenditure (Billions)",title = "Total NCAA Football Expenditure Over Time")+scale_x_continuous(breaks = c(2013,2014,2015,2016,2017,2018, 2019, 2020, 2021, 2022), minor_breaks = NULL)

# Top ten spenders and fpi teams
## find top 10 FPI 2023
df_23<-df[df$year==2022,]
df_sorted <- df_23[order(-df_23$fpi), ]

# Select the top 10 rows
top_10_colleges <- df_sorted[1:10, ]

plot1<-ggplot(top_10_colleges, aes(x=reorder(Data,fpi), y=fpi))+geom_bar(stat = "identity", position = position_dodge())+theme_minimal()+
  scale_x_discrete(labels = label_wrap(10))+
  labs(x="NCAA Football Team", y="FPI",title="Top 10 FPI Teams 2022")

## find top 10 expenses
df_23<-df[df$year==2022,]
df_sorted <- df_23[order(-df_23$Total_Expenses), ]

# Select the top 10 rows
top_10_colleges <- df_sorted[1:10, ]

plot2<-ggplot(top_10_colleges, aes(x=reorder(Data,Total_Expenses), y=Total_Expenses))+geom_bar(stat = "identity", position = position_dodge())+theme_minimal()+
  scale_x_discrete(labels = label_wrap(10))+
  labs(x="NCAA Football Team", y="Total Expenditure",title="Top 10 Expenditure Teams 2022")

stacked_plot <- grid.arrange(plot1, plot2, nrow = 2)
stacked_plot
```

### Line Plots
```{r}
# Conferences
df_plots <- df %>% left_join(Conference_key)

# create Date format year variable
df_plots <- df_plots %>%
  mutate(year = as.Date(paste0(year, "-01-01")))
df_plots <- df_plots %>%
  mutate(year = as.Date(paste0(year, "-01-01"), format = "%Y-%m-%d"))
df_plots$year <- format(df_plots$year, "%Y")

FBS <- c("Southeastern Conference", "Pacific-12 Conference",       
         "Atlantic Coast Conference", "Big Ten Conference","Big 12 Conference")
plots <- list()
for (i in 1:length(FBS)){
  dat <- df_plots %>% filter(FBS_Conference == FBS[i])
  
  plots[[i]] <- ggplot(dat, aes(x = year, y = fpi, group = Data, color = Data)) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(title = paste0(FBS[i], " FPI 2013 - 2022"), y = "FPI")
}
```


## Ten Year by Year LASSO Models
```{r}
library(tidyverse)
library(glmnet)

expenses <- c( "Other_Expenses", "Excess_Transfers_Back", "Medical", 
               "Competition_Guarantees_spent", "Recruiting", "Game_Expenses_and_Travel", 
               "Facilities_and_Equipment", "Coaches_Compensation", 
               "Support_and_Admin_Compensation_w_Severance", "Athletic_Student_Aid", 
               "Total_Football_Spending", "Total_Football_Coaching_Salaries" )
revenues <- c( "Corporate_Sponsorship_Advertising_Licensing", 
               "Donor_Contributions", "Competition_Guarantees_earned", 
               "NCAA_Conference_Distributions_Media_Rights_and_Post_Season_Football", 
               "Ticket_Sales", "Institutional_Government_Support")
# metrics not including fpi
metrics <- c("rk", "eff_ove","rnk_ove","eff_off","rnk_off","eff_def",
             "rnk_def","eff_spe","rnk_spe")

df <- read_csv("df_03202024.csv") %>%
  dplyr::select(Data, IPEDS_ID, year, fpi, Total_Expenses, all_of(expenses), Total_Revenues, all_of(revenues)) %>%
  filter(IPEDS_ID != 215293)

# Pivot the data wider - fpi
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year")
pivot_cols <- setdiff(colnames(df), id_cols)

df_wide <- df %>%
  pivot_wider(names_from = year,
              values_from = pivot_cols) %>%
  dplyr::select(-where(~all(is.na(.))))
```

### 2013 - 2014
```{r}
###### 2013 - 2014 -----------------------------------------------------------------
# all variables in 2013
set.seed(0619)
df_2013 <- df_wide %>%
  dplyr::select(contains(c("2013", "fpi_2014"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2013[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

### Lasso regression 2013 - 2014 - FPI -----------------------------
# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2014")
var_of_interest <- setdiff(colnames(df_2013), id_cols)

predictors <- scale(as.matrix(df_2013 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2013$fpi_2014


cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2014 - 2015
```{r}
###### 2014 - 2015 -----------------------------------------------------------------
# all variables in 2014
df_2014 <- df_wide %>%
  dplyr::select(contains(c("2014", "fpi_2015"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2014[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2015")
var_of_interest <- setdiff(colnames(df_2014), id_cols)

predictors <- scale(as.matrix(df_2014 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2014$fpi_2015


cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2015 - 2016 
```{r}
# all variables in 2015
df_2015 <- df_wide %>%
  dplyr::select(contains(c("2015", "fpi_2016"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2015[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2016")
var_of_interest <- setdiff(colnames(df_2015), id_cols)

predictors <- scale(as.matrix(df_2015 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2015$fpi_2016


cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2016 - 2017
```{r}
###### 2016 - 2017 -----------------------------------------------------------------
# all variables in 2016
df_2016 <- df_wide %>%
  dplyr::select(contains(c("2016", "fpi_2017"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2016[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2017")
var_of_interest <- setdiff(colnames(df_2016), id_cols)

predictors <- scale(as.matrix(df_2016 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2016$fpi_2017


cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2017 - 2018
```{r}
###### 2017 - 2018 -----------------------------------------------------------------
# all variables in 2017
df_2017 <- df_wide %>%
  dplyr::select(contains(c("2017", "fpi_2018"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2017[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2018")
var_of_interest <- setdiff(colnames(df_2017), id_cols)

predictors <- scale(as.matrix(df_2017 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2017$fpi_2018

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2018-2019
```{r}
# all variables in 2018
df_2018 <- df_wide %>%
  dplyr::select(contains(c("2018", "fpi_2019"))) %>%
  filter(complete.cases(.))

# Correlation matrix 
selected_columns <- df_2018[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2019")
var_of_interest <- setdiff(colnames(df_2018), id_cols)

predictors <- scale(as.matrix(df_2018 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2018$fpi_2019

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2019 - 2020
```{r}
###### 2019 - 2020 -----------------------------------------------------------------
# all variables in 2019
df_2019 <- df_wide %>%
  dplyr::select(contains(c("2019", "fpi_2020"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2019[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2020")
var_of_interest <- setdiff(colnames(df_2019), id_cols)

predictors <- scale(as.matrix(df_2019 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2019$fpi_2020

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2020 - 2021
```{r}
# all variables in 2020
df_2020 <- df_wide %>%
  dplyr::select(contains(c("2020", "fpi_2021"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2020[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2021")
var_of_interest <- setdiff(colnames(df_2020), id_cols)

predictors <- scale(as.matrix(df_2020 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2020$fpi_2021

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2021 - 2022 
```{r}
# all variables in 2021
df_2021 <- df_wide %>%
  dplyr::select(contains(c("2021", "fpi_2022"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2021[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2022")
var_of_interest <- setdiff(colnames(df_2021), id_cols)

predictors <- scale(as.matrix(df_2021 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2021$fpi_2022

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### 2022 - 2023
```{r}
# all variables in 2022
df_2022 <- df_wide %>%
  dplyr::select(contains(c("2022", "fpi_2023"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

# Correlation matrix 
selected_columns <- df_2022[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("Data", "IPEDS_ID", "FBS_Conference", "year", "fpi_2023")
var_of_interest <- setdiff(colnames(df_2022), id_cols)

predictors <- scale(as.matrix(df_2022 %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- df_2022$fpi_2023

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

## Final Generalized Lasso Model Process
```{r}
raw_data<-read.csv("df_03122024.csv")

vars_of_interest <- raw_data %>%
  rename(School=Data)%>%
  select(School, fpi, Recruiting,Total_Football_Spending,
         Total_Football_Coaching_Salaries,Institutional_Government_Support,
         Competition_Guarantees_earned,Competition_Guarantees_spent,
         Donor_Contributions,Other_Expenses,Medical,Ticket_Sales,
         Athletic_Student_Aid,Coaches_Compensation,Excess_Transfers_Back,
         NCAA_Conference_Distributions_Media_Rights_and_Post.Season_Football,
         Game_Expenses_and_Travel) %>%
  filter_all(all_vars(!is.na(.)))

averaged_data2 <- vars_of_interest %>%
  group_by(School) %>%
  summarize(
    fpi_avg = mean(fpi, na.rm=TRUE),
    Recruiting_avg = mean(Recruiting, na.rm = TRUE),
    Total_Football_Spending_avg = mean(Total_Football_Spending, na.rm = TRUE),
    Total_Football_Coaching_Salaries_avg = mean(Total_Football_Coaching_Salaries, na.rm = TRUE),
    Institutional_Government_Support_avg = mean(Institutional_Government_Support, na.rm = TRUE),
    Competition_Guarantees_earned_avg = mean(Competition_Guarantees_earned, na.rm = TRUE),
    Competition_Guarantees_spent_avg = mean(Competition_Guarantees_spent, na.rm = TRUE),
    Donor_Contributions_avg = mean(Donor_Contributions, na.rm = TRUE),
    Other_Expenses_avg = mean(Other_Expenses, na.rm = TRUE),
    Medical_avg = mean(Medical, na.rm = TRUE),
    Ticket_Sales_avg = mean(Ticket_Sales, na.rm = TRUE),
    Athletic_Student_Aid_avg = mean(Athletic_Student_Aid, na.rm = TRUE),
    Coaches_Compensation_avg = mean(Coaches_Compensation, na.rm = TRUE),
    Excess_Transfers_Back_avg = mean(Excess_Transfers_Back, na.rm = TRUE),
    NCAA_Conference_Distributions_Media_Rights_and_Post.Season_Football_avg = mean(NCAA_Conference_Distributions_Media_Rights_and_Post.Season_Football, na.rm = TRUE),
    Game_Expenses_and_Travel_avg = mean(Game_Expenses_and_Travel, na.rm = TRUE),
  
  )


df_041124 <- df_wide %>%
  dplyr::select(contains(c("Data","fpi_2023"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

end_df <- merge(averaged_data2, df_041124, by.x = "School", by.y = "Data", all.x = TRUE) %>% 
  select(-"School")

# Correlation matrix 
selected_columns <- end_df[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("fpi_2023")
var_of_interest <- setdiff(colnames(end_df), id_cols)

predictors <- scale(as.matrix(end_df %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- end_df$fpi_2023

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### LASSO on Predictors where MSE < 30
```{r}
raw_data<-read.csv("df_03122024.csv")

vars_of_interest <- df %>%
  rename(School=Data)%>%
  select(School, fpi, Recruiting,Total_Football_Spending,
         Total_Football_Coaching_Salaries,Institutional_Government_Support,
         Competition_Guarantees_earned, Competition_Guarantees_spent,
         Other_Expenses, Coaches_Compensation, Excess_Transfers_Back,
         Game_Expenses_and_Travel) %>%
  filter_all(all_vars(!is.na(.)))

averaged_data2 <- vars_of_interest %>%
  group_by(School) %>%
  summarize(
    fpi_avg = mean(fpi, na.rm=TRUE),
    Recruiting_avg = mean(Recruiting, na.rm = TRUE),
    Total_Football_Spending_avg = mean(Total_Football_Spending, na.rm = TRUE),
    Total_Football_Coaching_Salaries_avg = mean(Total_Football_Coaching_Salaries, na.rm = TRUE),
    Institutional_Government_Support_avg = mean(Institutional_Government_Support, na.rm = TRUE),
    Competition_Guarantees_earned_avg = mean(Competition_Guarantees_earned, na.rm = TRUE),
    Competition_Guarantees_spent_avg = mean(Competition_Guarantees_spent, na.rm = TRUE),
    Other_Expenses_avg = mean(Other_Expenses, na.rm = TRUE),
    Coaches_Compensation_avg = mean(Coaches_Compensation, na.rm = TRUE),
    Excess_Transfers_Back_avg = mean(Excess_Transfers_Back, na.rm = TRUE),
    Game_Expenses_and_Travel_avg = mean(Game_Expenses_and_Travel, na.rm = TRUE),
  )


df_041124 <- df_wide %>%
  dplyr::select(contains(c("Data","fpi_2023"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

end_df <- merge(averaged_data2, df_041124, by.x = "School", by.y = "Data", all.x = TRUE) %>% 
  select(-"School")

# Correlation matrix 
selected_columns <- end_df[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("fpi_2023")
var_of_interest <- setdiff(colnames(end_df), id_cols)

predictors <- scale(as.matrix(end_df %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- end_df$fpi_2023

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### Predictors with MSE < 30 and Avg LASSO Coeff was Highest
```{r}
vars_of_interest <- df %>%
  rename(School=Data)%>%
  select(School, fpi, Recruiting,Institutional_Government_Support,
         Other_Expenses) %>%
  filter_all(all_vars(!is.na(.)))

averaged_data2 <- vars_of_interest %>%
  group_by(School) %>%
  summarize(
    fpi_avg = mean(fpi, na.rm=TRUE),
    Recruiting_avg = mean(Recruiting, na.rm = TRUE),
    Institutional_Government_Support_avg = mean(Institutional_Government_Support, na.rm = TRUE),
    Other_Expenses_avg = mean(Other_Expenses, na.rm = TRUE),
  )


df_041124 <- df_wide %>%
  dplyr::select(contains(c("Data","fpi_2023"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

end_df <- merge(averaged_data2, df_041124, by.x = "School", by.y = "Data", all.x = TRUE) %>% 
  select(-"School")

# Correlation matrix 
selected_columns <- end_df[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("fpi_2023")
var_of_interest <- setdiff(colnames(end_df), id_cols)

predictors <- scale(as.matrix(end_df %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- end_df$fpi_2023

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```

### Using Only Significant Predictors
```{r}
set.seed(413619)
vars_of_interest <- df %>%
  rename(School=Data)%>%
  select(School, fpi, Recruiting,Total_Football_Spending,
         Total_Football_Coaching_Salaries,Institutional_Government_Support,
         Competition_Guarantees_earned,Competition_Guarantees_spent,
         Donor_Contributions,Other_Expenses,Medical,Ticket_Sales,
         Athletic_Student_Aid,Coaches_Compensation,Excess_Transfers_Back,
         NCAA_Conference_Distributions_Media_Rights_and_Post_Season_Football,
         Game_Expenses_and_Travel) %>%
  filter_all(all_vars(!is.na(.)))

averaged_data2 <- vars_of_interest %>%
  group_by(School) %>%
  summarize(
    fpi_avg = mean(fpi, na.rm=TRUE),
    Recruiting_avg = mean(Recruiting, na.rm = TRUE),
    Donor_Contributions_avg = mean(Donor_Contributions, na.rm = TRUE),
    Other_Expenses_avg = mean(Other_Expenses, na.rm = TRUE),
    Ticket_Sales_avg = mean(Ticket_Sales, na.rm = TRUE),
  )


df_041124 <- df_wide %>%
  dplyr::select(contains(c("Data","fpi_2023"))) %>%
  dplyr::select(-where(~ all(is.na(.) | . == 0)))

end_df <- merge(averaged_data2, df_041124, by.x = "School", by.y = "Data", all.x = TRUE) %>% 
  select(-"School")

# Correlation matrix 
selected_columns <- end_df[,]
correlation_matrix <- cor(selected_columns, use = "complete.obs") 
correlation_df <- as.data.frame(correlation_matrix)

# Preprocess the predictor variables (standardization)
id_cols <- c("fpi_2023")
var_of_interest <- setdiff(colnames(end_df), id_cols)

predictors <- scale(as.matrix(end_df %>% dplyr::select(var_of_interest)))
# Extract the response variable
fpi <- end_df$fpi_2023

cv <- cv.glmnet(x = predictors, y = fpi, 
                alpha = 1, 
                nfolds = 10 )

optimal_lambda <- cv$lambda.min

lasso_model <- glmnet(x = predictors, 
                      y = fpi, 
                      alpha = 1, 
                      lambda = optimal_lambda)
coef(lasso_model)

predictions <- predict(lasso_model, newx = predictors)

# Calculate performance metrics
mse <- mean((fpi - predictions)^2)
mse
rmse <- sqrt(mse)
rmse
```














